Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-05-07 doc/GMP2015/revision_2/appendix_defs.tex
Major changes to paper (revision 2)

\section{Basic Definitions and Theory}
\label{sec:appendix:defs}

In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
%
When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where $G$ is the underlying constraint (hyper)graph $G = (V,E)$ with the vertices $V$ representing the geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$ representing the constraints, each with an associated parameter $\delta$.
%
For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.

% \noindent
% \note Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.

\sidenote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}

In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}). In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).

Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values). Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates). For example, the \dfn{bar-joint rigidity matrix of the graph $G = (V,E)$} is a matrix of indeterminates representing the Jacobian of the distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.

% XXXXXXXXX
% something has to be said along with asimow roth
% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
% appropriate rank, the realizations or solutions of the corresponding
% constraint system are generically isolated and zero-dimensional if the
% constraint system has equalities and exact values for the parameters
% $\delta$. If the constraint system has inequalities, i.e, if the
% $\delta$ lie in an interval, this claim is approximate: the solutions
% are isolated small, full-dimensional  connected components.

One important property of generic constraint system or (hyper)graph
%that depends on the appropriate rigidity matrix of indeterminates
is \dfn{rigidity} (\note we refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably), i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
%This condition is called generic {\em infinitesimal rigidity}.

% \noindent
% \note If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.

\sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system

Other generic constraint system or (hyper)graph properties are the following:
% (\note we refer to these as properties of the constraint
% system or as properties of the underlying (hyper)graph
% interchangeably).
A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
%It is {\em rigid} if the number of independent rows of the rigidity matrix is
%at least the number of columns less the number of rigid body motions,
%which is 3 for distance constraint systems.
It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it is independent and not rigid, or \dfn{overconstrained} if it is not independent.

Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs. For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane. For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition. Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.

% In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.

% % \medskip\noindent
% % \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.

% \sidenote{In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications (\ref{sec:bodypin} and \ref{sec:pinnedline}).}

%XXXNow comes the definition of stress vector and flex vector.
%NOTE: something has to be said about what flexes and stresses mean

%TODO
Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}. It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions). For rigid graphs, all flex vectors are identically zero.
%
A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}. It has one entry per edge and represents a \dfn{self stress} of the system. For independent graphs, all stress vectors are identically zero.
%
An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.

>>>>

\section{Basic Definitions and Theory}
\label{sec:appendix:defs}
% \label{sec:prelim:defs}

Moved into prelims section.

### 2015-05-07 doc/GMP2015/revision_2/s01_introduction.tex
Major changes to paper (revision 2)

\todo{Main overview/intro.}
Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields. However, the design and analysis of these systems is made difficult by i

\subsection{Recursive Decomposition Problem}
The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph. We call these systems \nameit. This characterization has broad applicability in practice, and we provide several concrete examples.

Such decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads. In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which, in general, is NP-hard. Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan. By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan; we show that all canonical DR-plans are optimal for systems that are not overconstrained. This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties. In fact, we conjecture that one of the earlier algorithms also produces an optimal DR-plan and that every optimal DR-plan is in some formal sense canonical.

@@

When realizing such a system \nameit, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints). Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified). With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan. Thus, finding an optimal DR-plan is important for realization.

\subsection{Recombination Problem}
\todo{The second problem we consider...}
We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination. Once a subsystem is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.

@@

% \subsection{old}

@@

\todo{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan. This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}

@@

In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}. Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components. When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

\todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification. By dropping edges to get a convex configuration space, realizations of the original linkage can be found. Also problem relations...}

@@

We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.

\subsection{The Importance of an Optimal DR-plan}
% \subsection{Importance of an Optimal DR-plan}
% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.

Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan. Each block is the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.

>>>>

% \todo{Regarding qusecs... We use the standard way of dealing w/ these problems, namely modeling it as a geometric constraint system...
% \todo{Restructure intro so that we have... problems first... closing with ``Concerning these issue... this is what we do in relation to them... the specifics we address..'' and go into summary of our contributions}

Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields. Some systems have an underlying sparsity condition that characterize rigidity, allowing for matroidal techniques of rigidity detection and decomposition. However, the design and analysis of these systems is complicated by the NP-hardness barrier of creating optimal decomposition-recombination (DR-) plans. While other papers have focused on a few desirable criteria and made use of hueristics, by focusing on the class of independent (isostatic or underconstrained) systems we have found an efficient way to construct the optimal DR-plan.
Additionally, in real world applications, many of these systems exhibit degrees of self-similarity and quasi-uniformity. Our formulation of the optimal DR-plan can leverage these properties, further reducing the complexity of the DR-plan.
Furthermore, we address the fact that indecomposable subsystems in a system can be arbitrarily large, representing a bottle-neck in the computational efficiency in the analysis of these systems. We also address this problem with an efficient technique of optimal modification for further decomposition.

% In this paper, we introduce new techniques for the recursive decomposition and recombination of two dimensional independent (isostatic or underconstrained) constraint systems. The main contributions are a method to generate an optimal decomposition-recombination (DR-) plan

%  leveraging the quasi-uniformity and self-similarity present in many real world applications,

\subsection{Recursive Decomposition Problem}
The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph. This characterization has broad applicability in practice, and we provide several concrete examples.

% Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads. When realizing these sparsity-condition constraint systems, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints). Without decomposition to plan the recombination (DR-planning), this requires double exponential time in the number of variables (even if orientation type is specified). With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan. Thus, finding an optimal DR-plan is important for realization.

Decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads. When realizing constraint systems, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints); this requires double exponential time in the number of variable, even if orientation type is specified.
% Without decomposition to plan the recombination (DR-planning), this requires double exponential time in the number of variables (even if orientation type is specified).
With the use of decomposition-recomination (DR-) planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan. Thus, finding an optimal DR-plan is important for realization.

% \subsection{The Importance of an Optimal DR-plan}
% \subsection{Importance of an Optimal DR-plan}
% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.

% Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan. Each block is the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.

Furthermore, the DR-plan can be used to obtain recursive block decomposition of the rigidity matrix, stress space, and external stresses. Each block is the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of the system is given, this can be used to analyze the distribution of a given external load.

% In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which,
An optimal DR-plan decreases the computation involved in such tasks; however, in general, finding an optimal DR-plan is NP-hard. Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan. By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan. Then we show that all canonical DR-plans are optimal for systems that are not overconstrained. This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties. In fact, we conjecture that one of the earlier algorithms also produces an optimal DR-plan and that every optimal DR-plan is in some formal sense canonical.

@@

\subsection{Recombination Problem}
The second problem we consider is the complexity of recombination. Na{\"i}vely, recombination can be done by finding the solutions to the polynomial system of equations describing the constraints between the children; however, even in an optimal DR-plan, there can be arbitrarily many children at a node.

% Instead, we propose the modification of the subsystem to find a convex configuration space that can be efficiently searched for a solution to the original system.

Instead, we make use of the notion of Cayley complexity to quantify the difficulty of each step of recombination. Once a subsystem of a node is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.

% We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination. Once a subsystem is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
% This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.

@@

@@

In Section \ref{sec:prelim}, we provide basic definitions in rigidity theory, including what it means for a DR-plan to be complete or optimal. We also elaborate on the new notion of qusecs.
This section is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts. \todo{We don't model them as bar-joint... What did this mean originally?}

@@

In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}. Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner.
%
When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).
%
In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components.

% \todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification. By dropping edges to get a convex configuration space, realizations of the original linkage can be found. Also problem relations...}

@@

\todo{We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.}
